Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"

### 2019-11-12 writing/ehp_results.tex
Work some more on draft

We explored potential limitations in these data by comparing them with data
from other available sources.  For estimates of storm-associated rainfall, we
compared estimates used in the open-source data with ground-based observations
in nine sample counties (Figure~\ref{fig:raincomparison}). Within these
counties, storm-related rainfall measurements were well-correlated between the
two data sources, with rank correlations (bottom right of each graph in
Figure~\ref{fig:raincomparison}) between 0.87 and 0.98. There was some evidence
that the reanalysis data included in the open-source dataset may tend to
underestimate rainfall totals in storms with extremely high rainfall, based on
a few heavy-rainfall storms in Harris County, TX, Mobile County, AL, Charleston
County, SC, and Wake County, NC (Figure~\ref{fig:raincomparison}). However,
even in these cases, it was rare for a storm to be classified differently
(under the precipitation threshold of 75~\si{\milli\metre} we used for binary
exposure classification for later analysis) based on the source of
precipitation data---horizontal and vertical lines in each small plot in
Figure~\ref{fig:raincomparison} show the threshold of 75~\si{\milli\metre}, so
storms in the lower left and upper right quadrants would be classified the same
(``exposed'' or ``unexposed'') regardless of the precipitation data source,
while storms in the upper left and lower right quadrants would be classified
differently. Such cases are rare.

@@

We used this storm exposure data to classify counties as exposed or unexposed
to four different hazards for each tropical cyclone and then explored patterns
over years with available data. Across the four storm hazards considered, there
was wide variation in the average number of county exposures per year
(Table~\ref{tab:exposuresummaries}). For the tropical cyclone tornado exposure,
there were on average about 50 county exposures per year within our study.
County exposures were more frequent for tropical cyclone wind exposures ($>$140
/ yr on average), even more frequent for tropical cyclone flood exposures
($>$200 / yr on average), and most frequent for tropical cyclone rain exposure
($>$400 / yr on average). For every hazard except tornadoes, we identified at
least one tropical cyclone that exposed over~250 counties
(Table~\ref{tab:exposuresummaries}).  However, the largest-extent tropical
cyclone varied across hazards: Frances in~2004 exposed the most counties based
on rain, Ike in~2008 based on wind, and Ivan in 2004 based on flooding and
tornadoes (Table~\ref{tab:exposuresummaries}).

When we calculated and mapped the average number of exposures per decade in
each county for single-hazard exposures (Figure~\ref{fig:averageexposure}),
strong geographical patterns were clear. Tropical cyclone wind exposure had a
strong coastal pattern, with almost all exposures in counties within about~200
\si{\kilo\metre} (124~mi) of the coastline. While tropical cyclone rain
exposure was also more frequent in coastal areas compared to inland areas,
there were also inland exposures in counties that were never classified as
exposed based on winds. Flood-based exposures were frequently in the
Mid-Atlantic region, with a pattern that skewed north compared with other
exposures.Rain and, to some extent, flood exposures were characterized by a
pattern defined by the Appalachian Mountains, with fewer exposures west of the
mountain range than to the east. Almost all tornado-based exposures were in
coastal states, with many in Florida, and almost none north of Maryland.
Patterns were similar when analysis was restricted to years with exposure data
for all four hazards available (1996\,--\,2011; Figure~S3).

@@

Next, we measured the agreement between exposure classifications for each pair
of hazards, as well as between each hazard and exposure assessment based on
distance from the storm's track. We found that these exposure classifications
typically did not strongly agree between pairs of metrics---the set of counties
identified as ``exposed'' based on one metric often overlapped little with the set
identified as ``exposed'' by another metric, as storms frequently brought
different hazards to different locations.

@@

We drew similar conclusions when we investigated all~46 tropical cyclones
between~1996 and~2011 (the period for which data for all five metrics were
available) for which~100 or more counties were exposed based on at least one
metric (Figure~\ref{fig:jaccard}). In this figure, each row provides results
for one tropical cyclone, and each box in that row shows the measured Jaccard
coefficient for a pair of metrics.  For almost all tropical cyclones, Jaccard
index values between distance and any of the four hazard-based metrics
were~$\le0.6$ (i.e., out of the counties assessed as exposed by at least one of
the two metrics in the pair,~60\si{\percent} or fewer were assessed the same
under both metrics in the pair). For many tropical cyclones, the Jaccard index
values were~$\le0.5$, indicating at best moderate agreement.  Among the four
hazard-based metrics, the tornado-based metric had universally poor agreement
with other metrics in county-level classification across the tropical cyclones
considered.  Agreement between other pairs of metrics were better, but still
typically at best moderate (Jaccard index values usually $\le0.5$).

There were, however, exceptions\,---\,tropical cyclones in which exposure
assessment agreed well across several of the metrics considered.  For Floyd
in~1999 (Figure~S4) and Irene in~2011, for example, county-level classification
agreed moderately to well for all pairs of exposure metrics except those
involving the tornado-based metric (Figure~\ref{fig:jaccard}). For another set
of tropical cyclones (e.g., Ernesto in~2006, and Bertha in~1996, Isabel
in~2003), there was moderate to good agreement for all pairwise combinations of
distance, rain, and wind, but poor agreement for other combinations of
metrics, while for another set of storms (e.g., Matthew in~2004 and Katrina
in~2005), there was moderate to good agreement between distance and rain.
However, these sets of tropical cyclones represented the minority of tropical
cyclones considered; for most tropical cyclones, there was low overlap in the
counties determined to be exposed to the tropical cyclones between most
pairings of metrics.

>>>>

We explored potential limitations in these data by comparing them with data
from other available sources.  For estimates of storm-associated rainfall, we
compared estimates included in the open-source data with ground-based
observations in nine sample counties (Figure~\ref{fig:raincomparison}). Within
these counties, storm-related rainfall measurements were well-correlated
between the two data sources, with rank correlations (bottom right of each
graph in Figure~\ref{fig:raincomparison}) between 0.87 and 0.98. There was some
evidence that the primary rainfall metric may tend to underestimate rainfall
totals in storms with extremely high rainfall, based on a few heavy-rainfall
storms in Harris County, TX, Mobile County, AL, Charleston County, SC, and Wake
County, NC (Figure~\ref{fig:raincomparison}). However, even in these cases, it
was rare for a storm to be classified differently (under the precipitation
threshold of 75~\si{\milli\metre} we used for binary exposure classification
for later analysis) based on the source of precipitation data---horizontal and
vertical lines in each small plot in Figure~\ref{fig:raincomparison} show the
threshold of 75~\si{\milli\metre}, so storms in the lower left and upper right
quadrants would be classified the same (``exposed'' or ``unexposed'')
regardless of the precipitation data source, while storms in the upper left and
lower right quadrants would be classified differently. Such cases were rare.

@@

We used this storm exposure data to classify counties as exposed or unexposed
to four different hazards for each tropical cyclone and then explored patterns
over years with available data. Across the four storm hazards considered, there
was wide variation in the average number of county exposures per year
(Table~\ref{tab:exposuresummaries}). For tropical cyclone tornadoes,
there were on average about 50 county exposures per year within our study.
County exposures were more frequent for tropical cyclone wind exposures ($>$140
/ yr on average), even more frequent for tropical cyclone flood exposures
($>$200 / yr on average), and most frequent for tropical cyclone rain exposure
($>$400 / yr on average). For every hazard except tornadoes, we identified at
least one tropical cyclone that exposed over~250 counties
(Table~\ref{tab:exposuresummaries}).  However, the largest-extent tropical
cyclone varied across hazards: Frances in~2004 exposed the most counties based
on rain, Ike in~2008 based on wind, and Ivan in 2004 based on flooding and
tornadoes (Table~\ref{tab:exposuresummaries}).

When we calculated and mapped the average number of exposures per decade in
each county for single-hazard exposures (Figure~\ref{fig:averageexposure}),
strong geographical patterns were clear. Tropical cyclone wind exposure had a
strong coastal pattern, with almost all exposures in counties within about~200
\si{\kilo\metre} (124~mi) of the coastline. While tropical cyclone rain
exposure was also more frequent in coastal areas compared to inland areas,
there were also inland rain exposures in counties that were never classified as
exposed based on winds. Flood-based exposures were frequently in the
Mid-Atlantic region, with a pattern that skewed north compared with other
exposures. Rain and, to some extent, flood exposures were characterized by a
pattern defined by the Appalachian Mountains, with fewer exposures west of the
mountain range than to the east. Almost all tornado-based exposures were in
coastal states, with many in Florida, and almost none north of Maryland.
Patterns were similar when analysis was restricted to years with exposure data
for all four hazards available (1996\,--\,2011; Figure~S3).

@@

Next, we measured the agreement between exposure classifications for each pair
of hazards, as well as between with a proxy measurement based on
distance from the storm's track. We found that these exposure classifications
typically did not agree strongly between pairs of metrics---the set of counties
identified as ``exposed'' based on one metric often overlapped little with the set
identified as ``exposed'' by another metric, as storms frequently brought
different hazards to different locations.

@@

We drew similar conclusions when we investigated all~46 tropical cyclones
between~1996 and~2011 (the period for which data for all five metrics were
available) for which~100 or more counties were exposed based on at least one
metric (Figure~\ref{fig:jaccard}). In this figure, each row provides results
for one tropical cyclone, and each box in that row shows the Jaccard
coefficient for a pair of metrics.  For almost all tropical cyclones, Jaccard
coefficients were~$\le0.6$ between the distance-based metric and any of the
four hazard-based metrics (i.e., out of the counties assessed as exposed by at
least one of the two metrics in the pair,~60\si{\percent} or fewer were
assessed the same under both metrics). For many tropical cyclones, Jaccard
coefficients were~$\le0.5$, indicating at best moderate agreement.  Among the
four hazard-based metrics, the tornado-based metric had universally poor
agreement with other metrics in county-level classification across the tropical
cyclones considered.  Agreement between other pairs of metrics were better, but
still typically at best moderate (Jaccard coefficients usually $\le0.5$).

There were, however, exceptions\,---\,tropical cyclones in which exposure
assessment agreed well across several of the metrics considered.  For Floyd
in~1999 (Figure~S4) and Irene in~2011, for example, county-level classification
agreed moderately to well for all pairs of exposure metrics except those
involving the tornado-based metric (Figure~\ref{fig:jaccard}). For another set
of tropical cyclones (e.g., Ernesto in~2006, and Bertha in~1996, Isabel
in~2003), there was moderate to good agreement for all pairwise combinations of
distance, rain, and wind, but poor agreement for other combinations of metrics,
while for another set of storms (e.g., Matthew in~2004 and Katrina in~2005),
there was moderate to good agreement between distance and rain.  However, these
sets of tropical cyclones represented the minority of tropical cyclones
considered; for most tropical cyclones, there was low to moderate overlap in
the counties determined to be exposed to the tropical cyclones between most
pairings of metrics.
